Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-21 PAPER.tex
read through up until 2.2, corrected minor syntax stuff and also added new summary bullet point

\documentclass[preprint2,numberedappendix,tighten]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
%\documentclass[twocolumn,numberedappendix]{emulateapj}
\shorttitle{Characterizing Signal Loss, Error, and Bias in the 21cm Reionization Power Spectrum}
\shortauthors{Cheng et al.}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. While we only focus on a subset of PAPER-64 data in this paper, revised power spectrum limits from the PAPER experiment are presented in a companion paper by Kolopanis et al. (\textit{in prep.}) and supersede results from 
previously published PAPER analyses.
\end{abstract}

@@

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. 21\,cm tomography offers a unique window into the 
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013b, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, error estimation, and bias) that are essential to investigate 
for a robust 21\,cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER. In this study we use a subset of the PAPER-64 data to illustrate our revised analysis methods, while a companion paper, Kolopanis et al. (\textit{in prep.}), builds off of the methods in this paper to present revised PAPER-64 results for multiple redshifts and baseline types.

@@

This paper is organized into two main halves. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al. (\textit{in prep.})). We conclude in Section \ref{sec:Con}.

@@

%A deep understanding of the following three themes is essential for the accuracy and interpretation of a 21\,cm power spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important checklist for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Signal loss refers to attenuation of the target \textit{cosmological} signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if the amount of loss is not quantified accurately, it could lead to false non-detections and overly aggressive upper limits. Determining whether an analysis pipeline is lossy, and estimating the amount of loss if so, has subtle challenges but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on a 21\,cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Additionally, accurate error estimation is crucial for the comparison of results to theoretical models. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving that a bias is an EoR detection may be the most difficult challenge for 21\,cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must 
pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general 21\,cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter averages visibilities in time to produce a smaller number of more sensitive 
independent samples. However, such a filter also affects the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by previous PAPER analyses is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). Using such a technique enables the down-weighting of contaminant modes that obey a different covariance structure from that of cosmological modes. However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two has large impacts on signal loss. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Namely, bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011b}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass both jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use a toy data model to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be performed for robust 21\,cm results.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as by fitting a polynomial during spectral calibration, applying a delay-domain filter, or deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
% JEA - we're not considering a general problem, but are really just focused on the empirical QE and its discontents, so let's say that
the use of an empirically estimated covariance matrix with the ``optimal quadratic estimator'' formalism.
This loss was significantly 
%applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously 
under-estimated in the \citetalias{ali_et_al2015} analysis.

@@

% JEA - The discussion of dimensions here is confusing ... "a baseline" is mentioned, but it's not clear if the redundant baseline averaging has already happened or what.  I believe we are only considering covariance between frequencies, but this isn't explicitly said, and the matrix equations don't make sense unless all the dimensions are N_f N_t.  
%We begin with our 
The measured visibilities for a single baseline in Jy are arranged as a data vector, $\textbf{x}$. It has length $N_{t} N_{f}$,
% JEA - not sure this helps
%, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), 
where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. The covariance of the data is given by 
\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle = \textbf{S} + \textbf{U}
\end{equation}
where the average over an ensemble of data realizations produces the true covariance, and we further assume it may be written as a sum of the desired cosmological signal $\textbf{S}$ and other terms $\textbf{U}$.

@@

The optimal quadratic estimator prescription is then to compute
\begin{equation}
\label{eq:OQE}
\widehat{p}^{\alpha}  = \sum_\beta ({\F^{-1}})^{\alpha\beta} (\widehat{q}^{\beta} - \widehat{b}^{\beta} )
\end{equation}
where $\F$ is the Fisher matrix 
\begin{equation}
F^{\alpha \beta} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \textbf{Q}^{\alpha} \C^{-1} \textbf{Q}^{\beta} \right)
\end{equation}
and
\begin{equation}
\label{eq:OQEQuadratic}
\widehat{q}^\alpha =  \half \textbf{x}^\dagger \invC \textbf{Q}^{\alpha}  \invC \textbf{x} 
\end{equation}
and
\begin{equation}
\label{eq:OQELinear}
\widehat{b}^{\alpha} = \half \Tr\left( \mathbf{U} \invC \textbf{Q}^{\alpha} \invC \right).
\end{equation}
This estimator is the minimum variance (smallest error bar) estimate of the power spectrum  subject to the constraint that it is also unbiased; that is, the ensemble average of the estimator is equal to its true value
\begin{equation}
\label{eq:super_unbiased}
\langle \widehat{p}^{\alpha} \rangle = p^\alpha.
\end{equation}
\citep{tegmark_et_al1997a,bond_et_al1998}

@@

%We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:
Motivated by the desire to retain the advantageous behavior of suppressing contributions of $\mathbf{U}$ to estimates of the EoR power spectrum, we note that is possible to define a modified version of the quadratic estimator 
%which may still produce an unbiased estimate, 
where Equation \eqref{eq:OQEQuadratic} is replaced by
\begin{equation}
\label{eq:qhat}
\widehat{q}^{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}
\end{equation}
where $\textbf{R}$ is a weighting matrix chosen by the data analyst.  For example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $\textbf{R} \equiv \textbf{I}$, the identity matrix.
The matrix $\textbf{Q}^{\alpha}$ encodes the dependence of the covariance on the power spectrum parameters $\partial \C/\partial p^\alpha$ but in practice also does other things, including implementing a transform of the frequency-domain visibilities to $\mathbf{k}$-space, taking into account cosmological scalings, and converting the visibilities from Jy to Kelvin.
%with dimensions ($N_{f}$,$N_{f}$) --- as

@@

% JEA - I don't think it's attractive at all
%Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself. 
Given the advantages of inverse covariance weighting, a question arises of how one goes about estimating $\C$.  One method is to empirically derive it from the data $\textbf{x}$ itself.
Similar types of weightings that are based on variance information in data are done in \citet{chang_et_al2010} and \citet{switzer_et_al2015}. In previous PAPER analyses, one time-averages the data to obtain:
\begin{equation}
\widehat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle_{t} \approx \langle  \textbf{xx}^{\dagger}\rangle,
\end{equation}
assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for our empirically estimated inverse covariance weighting is then $
\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$.

@@

\item If the covariance matrix is estimated from the data, the strong correlation between the estimated modes and the data will in general produce an estimate of the signal power spectrum which is strongly biased {\it low} relative to the true value.  In this context, this is what we call ``signal loss''  (Section \ref{sec:toymodel}).

@@

\item The convergence of the modes of the empirical covariance matrix to the true matrix depends on the size of the eigenvalues, with the modes with the largest eigenvalues converging more quickly (in norm) to the true eigenvector than those with smaller eigenvalues. Thus, in estimating the covariance matrix empirically, larger eigenvalue modes are likely to be more reliable (Section \ref{sec:toymodel_frf}).

\item Knowing these things, there are some simple ways of altering the empirical covariance matrix to produce unbiased power spectrum estimates (Section \ref{sec:otherweight}).

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. 
% JEA - redundant
%For an empirically estimated $\widehat{\textbf{C}}$, 
%its eigenvalues differ from those of the true covariance, where 
%the coupling between EoR-dominated eigenmodes and the data can lead to loss.
%Hence, 
There is negligible signal loss when all eigenmodes besides the foreground one are no longer correlated with the data.
%assigning identical weights to all eigenmodes except the first, 
%constructing a covariance proportional to the identity and the outer product of the sinusoid eigenmode, since we are only down-weighting the FG-dominated mode in this case.
}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

Armed with this information,
%Using what we've learned about the eigenspectrum, 
we can tweak the covariance in a simple way to suppress foregrounds and yield minimal signal loss. Recall that our toy model foreground 
% JEA - modes need not be sinusoids; the point is that there is one outer product describing this in the covariance
%is a sinusoid, so it 
can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the first eigenmode and 
%then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the EoR-dominating part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e. only allowing over-fitting to occur for the first mode).
then take the remaining covariance to be the identity matrix.  
This decouples the covariance from the data for the EoR modes.  The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. 
In this case we recover the EoR signal, demonstrating that if we can disentangle the foreground-dominated modes and EoR-dominated modes, we can suppress
% down-weight them 
foregrounds with negligible signal loss.

>>>>

\documentclass[preprint2,numberedappendix,tighten]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
%\documentclass[twocolumn,numberedappendix]{emulateapj}
\shorttitle{Characterizing Signal Loss, Error, and Bias in the 21\,cm Reionization Power Spectrum}
\shortauthors{Cheng et al.}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different analysis methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. While we only focus on a subset of PAPER-64 data in this paper, revised power spectrum limits from the PAPER experiment are presented in a companion paper by Kolopanis et al. (\textit{in prep.}) and supersede results from 
previously published PAPER analyses.
\end{abstract}

@@

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Hence, 21\,cm tomography offers a unique window into both the spatial and temporal
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013b, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, error estimation, and bias) that are essential to investigate 
for a robust 21\,cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER. In this study we use a subset of PAPER-64 data to illustrate our revised analysis methods, while a companion paper, Kolopanis et al. (\textit{in prep.}), builds off of the methods in this paper to present revised PAPER-64 results for multiple redshifts and baseline types.

@@

This paper is organized into two main halves. In Section \ref{sec:Themes} we introduce the three themes of our focus, using 
toy models to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the methods used in the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al. (\textit{in prep.})). We conclude in Section \ref{sec:Con}.

@@

%A deep understanding of the following three themes is essential for the accuracy and interpretation of a 21\,cm power spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important checklist for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Signal loss refers to attenuation of the target \textit{cosmological} signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if the amount of loss is not quantified accurately, it could lead to false non-detections and overly aggressive upper limits. Determining whether an analysis pipeline is lossy, and estimating the amount of loss if so, has subtle challenges but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on a 21\,cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Additionally, accurate error estimation is crucial for the comparison of results to theoretical models. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving that a bias is an EoR detection may be the most difficult challenge for future 21\,cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe example tests that a true EoR detection must 
pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general 21\,cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter averages visibilities in time to produce a smaller number of more sensitive 
independent samples. However, such a filter also affects the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by previous PAPER analyses is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). Using such a technique enables the down-weighting of contaminant modes that obey a different covariance structure from that of cosmological modes. However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two can have large impacts on signal loss. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Namely, bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011b}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass both jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use toy data models to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be performed for robust 21\,cm results.

@@

Signal loss can arise in a variety of ways in an analysis pipeline, such as by fitting a polynomial during spectral calibration, applying a delay domain filter, or deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
% JEA - we're not considering a general problem, but are really just focused on the empirical QE and its discontents, so let's say that
the use of an empirically estimated covariance matrix with the ``optimal quadratic estimator'' formalism.
This loss was significantly 
%applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously 
under-estimated in the \citetalias{ali_et_al2015} analysis.

@@

% JEA - The discussion of dimensions here is confusing ... "a baseline" is mentioned, but it's not clear if the redundant baseline averaging has already happened or what.  I believe we are only considering covariance between frequencies, but this isn't explicitly said, and the matrix equations don't make sense unless all the dimensions are N_f N_t.  
%We begin with our 
The measured visibilities for a single baseline in Jy are arranged as a data vector, $\textbf{x}$. It has length $N_{t} N_{f}$,
% JEA - not sure this helps
%, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), 
where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. The covariance of the data is given by 
\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle = \textbf{S} + \textbf{U}
\end{equation}
where the average over an ensemble of data realizations produces the true covariance, and we further assume it may be written as the sum of the desired cosmological signal $\textbf{S}$ and other terms $\textbf{U}$.

@@

The optimal quadratic estimator prescription is then to compute
\begin{equation}
\label{eq:OQE}
\widehat{p}^{\alpha}  = \sum_\beta ({\F^{-1}})^{\alpha\beta} (\widehat{q}^{\beta} - \widehat{b}^{\beta} )
\end{equation}
where $\F$ is the Fisher matrix 
\begin{equation}
F^{\alpha \beta} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \textbf{Q}^{\alpha} \C^{-1} \textbf{Q}^{\beta} \right)
\end{equation}
and
\begin{equation}
\label{eq:OQEQuadratic}
\widehat{q}^\alpha =  \half \textbf{x}^\dagger \invC \textbf{Q}^{\alpha}  \invC \textbf{x} 
\end{equation}
and
\begin{equation}
\label{eq:OQELinear}
\widehat{b}^{\alpha} = \half \Tr\left( \mathbf{U} \invC \textbf{Q}^{\alpha} \invC \right).
\end{equation}
This estimator is the minimum variance (smallest error bar) estimate of the power spectrum  subject to the constraint that it is also unbiased; that is, the ensemble average of the estimator is equal to its true value
\begin{equation}
\label{eq:super_unbiased}
\langle \widehat{p}^{\alpha} \rangle = p^\alpha
\end{equation}
\citep{tegmark_et_al1997a,bond_et_al1998}.

@@

%We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:
Motivated by the desire to retain the advantageous behavior of suppressing contributions of $\mathbf{U}$ to estimates of the EoR power spectrum, we note that is possible to define a modified version of the quadratic estimator 
%which may still produce an unbiased estimate, 
where Equation \eqref{eq:OQEQuadratic} is replaced by
\begin{equation}
\label{eq:qhat}
\widehat{q}^{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}
\end{equation}
where $\textbf{R}$ is a weighting matrix chosen by the data analyst.  For example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $\textbf{R} \equiv \textbf{I}$, the identity matrix.
The matrix $\textbf{Q}^{\alpha}$ encodes the dependence of the covariance on the power spectrum parameters $\partial \C/\partial p^\alpha$ but in practice also does other things, including implementing a transform of the frequency domain visibilities to $\mathbf{k}$-space, taking into account cosmological scalings, and converting the visibilities from Jy to Kelvin.
%with dimensions ($N_{f}$,$N_{f}$) --- as

@@

% JEA - I don't think it's attractive at all
%Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself. 
Given the advantages of inverse covariance weighting, a question arises of how one goes about estimating $\C$.  One method is to empirically derive it from the data $\textbf{x}$ itself.
Similar types of weightings that are based on variance information in data are done in \citet{chang_et_al2010} and \citet{switzer_et_al2015}. In previous PAPER analyses, one time-averages the data to obtain:
\begin{equation}
\widehat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle_{t} \approx \langle  \textbf{xx}^{\dagger}\rangle,
\end{equation}
assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for our empirically estimated inverse covariance weighting is then $
\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$, where we use a hat symbol to distinguish the empirical covariance from the true covariance $\textbf{C}$.

@@

\item If the covariance matrix is estimated from the data, a strong correlation between the estimated modes and the data will in general produce an estimate of the signal power spectrum which is strongly biased {\it low} relative to the true value.  In this context, this is what we call ``signal loss''  (Section \ref{sec:toymodel}).

@@

\item The rate at which empirical eigenvectors converge to their true forms depends on sample variance in the empirical estimate, the number of realizations in the estimate, and the steepness of the empirical eigenspectrum (Section \ref{sec:toymodel_frf}). In general, larger sample variances, fewer realizations, and flatter eigenspectra lead to more loss.

%\item The convergence of the modes of the empirical covariance matrix to the true matrix depends on the size of the eigenvalues, with the modes with the largest eigenvalues converging more quickly (in norm) to the true eigenvector than those with smaller eigenvalues. Thus, in estimating the covariance matrix empirically, larger eigenvalue modes are likely to be more reliable (Section \ref{sec:toymodel_frf}).

\item Knowing these things, there are some simple ways of altering the empirical covariance matrix to decouple it from the data and produce unbiased power spectrum estimates (Section \ref{sec:otherweight}).

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of the foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare using empirically estimated inverse covariance weighting where $\textbf{C}$ is derived from the data (left), and projecting out the zeroth eigenmode only (right). In the former case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. 
% JEA - redundant
%For an empirically estimated $\widehat{\textbf{C}}$, 
%its eigenvalues differ from those of the true covariance, where 
%the coupling between EoR-dominated eigenmodes and the data can lead to loss.
%Hence, 
There is negligible signal loss when all eigenmodes besides the foreground one are no longer correlated with the data.
%assigning identical weights to all eigenmodes except the first, 
%constructing a covariance proportional to the identity and the outer product of the sinusoid eigenmode, since we are only down-weighting the FG-dominated mode in this case.
}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

Armed with this information,
%Using what we've learned about the eigenspectrum, 
we can tweak the covariance in a simple way to suppress foregrounds and yield minimal signal loss. Recall that our toy model foreground 
% JEA - modes need not be sinusoids; the point is that there is one outer product describing this in the covariance
%is a sinusoid, so it 
can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the zeroth eigenmode and 
%then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the EoR-dominating part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e. only allowing over-fitting to occur for the first mode).
then take the remaining covariance to be the identity matrix.  
This decouples the covariance from the data for the EoR modes.  The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. 
In this case we recover the EoR signal, demonstrating that if we can disentangle the foreground-dominated modes and EoR-dominated modes, we can suppress
% down-weight them 
foregrounds with negligible signal loss.

Altering $\widehat{\textbf{C}}$ as such is one specific example of a regularization method for this toy model, in which we are changing $\widehat{\textbf{C}}$ in a way that changes its coupling to the data realization. There are several other simple ways to regularize $\widehat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.
